A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2015-09-11 docs/data_v1.01.tex
Removed some TO DOs.

\paragraph{Redshift.} [TO DO: WHO???] found from SDSS spectra that J1331 has two redshifts inside 1 arcsec: The smaller one, $z_d = 0.113$, is the redshift of J1331 itself, the larger one, $z_s = 0.254$, is the redshift of the lensed background source \citep{SWELLSIII}. According to its redshift and the WMAP5 [TO DO: CHECK] cosmology \citep{WMAP5cosm}, J1331 has an angular diameter distance of 414 Mpc, which translates into a transverse scaling of 1 arcsec $\hat{=}$ 2.01 kpc.

@@

\paragraph{Stellar kinematics.} For the dynamical modelling we also use the stellar kinematics for J1331 measured by \citet{SWELLSV}. They obtained long-slit spectra along J1331's major-axis with the Low Resolution Imaging Spectrograph (LRIS) on the Keck I 10m telescope. The width of the slot was 1 arcsec and the seeing Conditions had a FWHM of $\sim 1.1$ arcsec. Spectra for spatial bins of different widths along the major axis were extracted. Analogously to \citet{SWELLSII} they measured line-of-sight rotation velocities ($v_\text{rot}$) and stellar velocity dispersion ($\sigma$) by fitting Gaussian line profiles to emission lines in these spectra. Gas kinematics were extracted from fits to H$\alpha$ and NII lines, as tracers for inonized gas.
\\The stellar kinematics, $v_\text{rot}$, $\sigma$ and $v_\text{rms}^2=v_\text{rot}^2 + \sigma^2$ are shown in Figure \ref{fig:kinematics}. The rotation curve reveals a counter-rotating core within 2 arcsec $\simeq$ 4 kpc. Outside of $\sim 3.5 arcsec$ there is a steep drop in the dispersion, which is exptected at the boundary between the pressure supported bulge and the rotationally supported disk, which appears around this radius in the F450W filter in Figure \ref{fig:F450W}. However, in the brighter F814W filter in Figure \ref{fig:F814W}  the large reddish bulge extends out to $\sim$5 arcsec. 
\\Inside of $\sim 4$ arcsec, the data appears to be symmetric, outside of this the assumption of axisymmetry seems not to be valid anymore, considering the data. We add -2.3 km/s to the $v_\text{rot}$ to ensure $v_\text{rot}(R=0) \sim 0$ as a possible correction term for a systematic misjudgement of the systemic velocity. We also symmetrize the data within 4 arcsec and asign a minimum error of $\delta v_\text{rms} > 5$ km/s to the $v_\text{rms}$ data. In the JAM modelling, which is based on the assumption of axisymmetry, only kinematics within $\sim 2.5$ and 4 arcsec are used. Another reason to restrict to modelling on the bulge region is that our MGE in Table \ref{tab:MGEF814W} is only a good representation of J1331's F814W light distribution inside $\sim 5$ arcsec.\\

>>>>

\paragraph{Redshift.} [TO DO: WHO???] found from SDSS spectra that J1331 has two redshifts inside 1 arcsec: The smaller one, $z_d = 0.113$, is the redshift of J1331 itself, the larger one, $z_s = 0.254$, is the redshift of the lensed background source \citep{SWELLSIII}. According to its redshift and the WMAP5 [TO DO: Explain?] cosmology \citep{WMAP5cosm}, J1331 has an angular diameter distance of 414 Mpc, which translates into a transverse scaling of 1 arcsec $\hat{=}$ 2.01 kpc.

@@

\paragraph{Stellar kinematics.} For the dynamical modelling we also use the stellar kinematics for J1331 measured by \citet{SWELLSV}. They obtained long-slit spectra along J1331's major-axis with the Low Resolution Imaging Spectrograph (LRIS) on the Keck I 10m telescope. The width of the slot was 1 arcsec and the seeing Conditions had a FWHM of $\sim 1.1$ arcsec. Spectra for spatial bins of different widths along the major axis were extracted. Analogously to \citet{SWELLSII} they measured line-of-sight rotation velocities ($v_\text{rot}$) and stellar velocity dispersion ($\sigma$) by fitting Gaussian line profiles to emission lines in these spectra. Gas kinematics were extracted from fits to H$\alpha$ and NII lines, as tracers for inonized gas.
\\The stellar kinematics, $v_\text{rot}$, $\sigma$ and $v_\text{rms}^2=v_\text{rot}^2 + \sigma^2$ are shown in Figure \ref{fig:kinematics}. The rotation curve reveals a counter-rotating core within 2 arcsec $\simeq$ 4 kpc. Outside of $\sim 3.5$ arcsec there is a steep drop in the dispersion, which is exptected at the boundary between the pressure supported bulge and the rotationally supported disk, which appears around this radius in the F450W filter in Figure \ref{fig:F450W}. However, in the brighter F814W filter in Figure \ref{fig:F814W}  the large reddish bulge extends out to $\sim$5 arcsec. 
\\Inside of $\sim 4$ arcsec, the data appears to be symmetric, outside of this the assumption of axisymmetry seems not to be valid anymore, considering the data. We add -2.3 km/s to the $v_\text{rot}$ to ensure $v_\text{rot}(R'=0) \sim 0$ as a possible correction term for a systematic misjudgement of the systemic velocity. We also symmetrize the data within 4 arcsec and asign a minimum error of $\delta v_\text{rms} > 5$ km/s to the $v_\text{rms}$ data. In the JAM modelling, which is based on the assumption of axisymmetry, only kinematics within $\sim 2.5$ and 4 arcsec are used. Another reason to restrict to modelling on the bulge region is that our MGE in Table \ref{tab:MGEF814W} is only a good representation of J1331's F814W light distribution inside $\sim 5$ arcsec.\\

### 2015-09-11 docs/discussion_ML_v1.02.tex
Finished introduction (references still missing).
Removed some TO DOs.

An alternative JAM model was introduced in Section \ref{sec:results_JAM_SB}, where we used the surface brightness distribution without a halo component but together with an increasing mass-to-light ratio. We found that such a model would be perfectly consistent with the lensing results and predict a total mass-to-light ratio in the center of J1331 of  $\Upsilon_\text{I,tot}(R\sim0) = 2.53$. This is very close to the stellar mass-to-light ratio we got from \citet{SWELLSI} estimates of the bulge's stellar mass using the Chabrier IMF $\Upsilon_\text{I,*}^\text{Chabrier} = 2.5 \pm 0.6$, i.e. less bottom-heavy (more bluish) population.\\

@@

[TO DO: Einheitlich: $Upsilon_{I,*}^{comment}$, $\Upsilon_\text{I,tot}^\text{ein}$]

>>>>

An alternative JAM model was introduced in Section \ref{sec:results_JAM_SB}, where we used the surface brightness distribution without a halo component but together with an increasing mass-to-light ratio. We found that such a model would be perfectly consistent with the lensing results and predict a total mass-to-light ratio in the center of J1331 of  $\Upsilon_\text{I,tot}(R'\sim0) = 2.53$. This is very close to the stellar mass-to-light ratio we got from \citet{SWELLSI} estimates of the bulge's stellar mass using the Chabrier IMF $\Upsilon_\text{I,*}^\text{Chabrier} = 2.5 \pm 0.6$, i.e. less bottom-heavy (more bluish) population.\\

@@

[TO DO: Einheitlich: $\Upsilon_{I,*}^{comment}$, $\Upsilon_\text{I,tot}^\text{ein}$]

[TO DO: Why are disk and bulge mass in SWELLS I different to the one in SWELLS IV??]

### 2015-09-11 docs/introduction_v1.02.tex
Removed some TO DOs.

%Dark Matter
The flat rotation curves of galaxies (deduced from e.g. the stellar movements in the solar neighbourhood \citep{1932BAN.....6..249O} and $H\alpha$ rotation curves in external galaxies \citep{1978ApJ...225L.107R}) were the first indication that galaxies might reside in massive, almost spherical halos made of invisible ``dark'' matter. The standard model of cosmology, based on results by the Planck Mission [TO DO: REF], predicts that $\sim 85\%$ of the universe's total matter content is in the form of non-baryonic cold dark matter. Simulations suggest that this kind of dark matter forms cuspy halos following a Navarro-Frenck-White (NFW) profile \citep{NFW96}. However, central dark matter density cusps are not observed in dark matter dominated dwarf galaxies [TO DO: REF]. If they exist in more massive galaxies depends strongly on stellar mass-to-light ratio [TO DO: REF]. Overall, observations suggest cored dark matter halos. This discrepancy, known as the core-cusp problem, might be due to a yet unknown interaction between dark matter and baryons [TO DO: REF].\\

@@

\begin{figure*}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_450.ps}
  \caption{J1331 in F450W ("blue")}
  \label{fig:F450W}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_814.ps}
  \caption{J1331 in F814W ("red")}
  \label{fig:F814W}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/lens_imgpos.ps}
  \caption{The lensing images}
  \label{fig:lens_just_imgpos}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/stellar_kinematics_data.ps}
  \caption{Stellar Kinematics by \citet{SWELLSV}}
  \label{fig:kinematics}
\end{subfigure}
\caption{Hubble Space telescope (HST) images and stellar kinematics of the galaxy SDSS J1331+3638 (J1331), which has a large counter-rotating core and whose bulge acts as a strong lens for a bluish background source. \emph{Panel (a) and (b):} HST/WFPC2/WFC3 images of J1331 by \citet{SWELLSI} in two filters, F450W in panel (a) and F814W in panel (b). The galaxy's coordinates on the sky are right ascension $\alpha$ = 202.91800$^\circ$ and declination $\delta$ = 36.46999$^\circ$ (epoch J2000). Image orientation and scaling are indicated in panel (a); the scaling transformation from arcseconds to the physical size of the galaxy in kpc uses the galaxy's redshift $z_d = 0.113$ \citep{SWELLSIII} (i.e. assumes an angular diameter distance of 414 Mpc). The color scaling of these two images is the same. The black solid line in panel (b) shows the orientation of the major-axis. The line has a length of 10 arcsec and indicates the region within which we carry out the Jeans modelling. [TO DO: NOT ALL THE TIME.???????] \emph{Panel (c):} The central region of J1331 in F450W, surface brightness subtracted. An IRAF ellipse ???? fit to the F450W surface brightness in panel (a) was subtracted from the image. The (smoothed) residuals within the white square in panel (a) are shown in panel (c). Four bright blobs (A,B,C and D) become visible, which are arranged in a typical strong lensing configuration around the center of the galaxy (G). \emph{Panel (d):} Stellar Kinematics along the galaxy's major axis as measured by \citet{SWELLSV}, line-of-sight rotation velocity $v_\text{rot}$, line-of-sight velocity dispersion $\sigma$ and the rms-velocity $v_\text{rms} = \sqrt{v_\text{rot}^2 + \sigma^2}$. The dotted line in panel (b) indicates the galaxy's effective half-light radius (in the F814W filter), $R_\text{eff} = 2.6" = 5.2$ kpc. The $v_\text{rot}$ curve reveals that J1331 is counter-rotating within $R_\text{eff}$. [TO DO: Add (x,y) axis in figure b).???????]}
\label{fig:specialJ1331}
\end{figure*}

>>>>

%Dark Matter
The flat rotation curves of galaxies (deduced from e.g. the stellar movements in the solar neighbourhood \citep{1932BAN.....6..249O} and $H\alpha$ rotation curves in external galaxies \citep{1978ApJ...225L.107R}) were the first indication that galaxies might reside in massive, almost spherical halos made of invisible ``dark'' matter. The standard model of cosmology, based on results by the Planck Mission \citep{WMAP5cosm}, predicts that $\sim 85\%$ of the universe's total matter content is in the form of non-baryonic cold dark matter. Simulations suggest that this kind of dark matter forms cuspy halos following a Navarro-Frenck-White (NFW) profile \citep{NFW96}. However, central dark matter density cusps are not observed in dark matter dominated dwarf galaxies [TO DO: REF]. If they exist in more massive galaxies depends strongly on stellar mass-to-light ratio [TO DO: REF]. Overall, observations suggest cored dark matter halos. This discrepancy, known as the core-cusp problem, might be due to a yet unknown interaction between dark matter and baryons [TO DO: REF].\\

@@

\begin{figure*}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_450.ps}
  \caption{J1331 in F450W ("blue")}
  \label{fig:F450W}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_814.ps}
  \caption{J1331 in F814W ("red")}
  \label{fig:F814W}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/lens_imgpos.ps}
  \caption{The lensing images}
  \label{fig:lens_just_imgpos}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/stellar_kinematics_data.ps}
  \caption{Stellar Kinematics by \citet{SWELLSV}}
  \label{fig:kinematics}
\end{subfigure}
\caption{Hubble Space telescope (HST) images and stellar kinematics of the galaxy SDSS J1331+3638 (J1331), which has a large counter-rotating core and whose bulge acts as a strong lens for a bluish background source. \emph{Panel (a) and (b):} HST/WFPC2/WFC3 images of J1331 by \citet{SWELLSI} in two filters, F450W in panel (a) and F814W in panel (b). The galaxy's coordinates on the sky are right ascension $\alpha$ = 202.91800$^\circ$ and declination $\delta$ = 36.46999$^\circ$ (epoch J2000). Image orientation and scaling are indicated in panel (a); the scaling transformation from arcseconds to the physical size of the galaxy in kpc uses the galaxy's redshift $z_d = 0.113$ \citep{SWELLSIII} (i.e. assumes an angular diameter distance of 414 Mpc). The color scaling of these two images is the same. The black solid line in panel (b) shows the orientation of the major-axis. The line has a length of 10 arcsec and indicates the region within which we carry out the Jeans modelling. \emph{Panel (c):} The central region of J1331 in F450W, surface brightness subtracted. An IRAF ellipse [TO DO: How to write and reference????] fit to the F450W surface brightness in panel (a) was subtracted from the image. The (smoothed) residuals within the white square in panel (a) are shown in panel (c). Four bright blobs (A,B,C and D) become visible, which are arranged in a typical strong lensing configuration around the center of the galaxy (G). \emph{Panel (d):} Stellar Kinematics along the galaxy's major axis as measured by \citet{SWELLSV}, line-of-sight rotation velocity $v_\text{rot}$, line-of-sight velocity dispersion $\sigma$ and the rms-velocity $v_\text{rms} = \sqrt{v_\text{rot}^2 + \sigma^2}$. The dotted line in panel (b) indicates the galaxy's effective half-light radius (in the F814W filter), $R_\text{eff} = 2.6" = 5.2$ kpc. The $v_\text{rot}$ curve reveals that J1331 is counter-rotating within $R_\text{eff}$. [TO DO: Projected radius $R'$ on x-axis in panel d).] [TO DO: Add (x,y) axis in figure b).???????]}
\label{fig:specialJ1331}
\end{figure*}

### 2015-09-11 docs/model_JAM_v1.01.tex
Removed some TO DOs.

\\The JAM modelling approach by \citet{Cap08} makes use of expressing the tracer density and the mass density as MGEs (see also \citet{1994A&A...285..723E}). The density of stellar tracers is assumed to be proportional to the observed and deprojected brightness distribution $\nu(R,z)$ in Equation \ref{eq:deprojMGE}. The mass density $\rho(R,z)$ consists of several sets of MGEs: One MGE, that is usually taken to be $\nu(R,z)$ multiplied by a constant stellar mass-to-light ratio $\Upsilon_*$, describes the distribution of stellar mass in the galaxy. (To mimick gradients of mass-to-light ratio, each Gaussian could be assigned its own $\Upsilon_{*,i}$.) To add a Navarro-Frenck-White (NFW) \citep{Navarro+1995c,NFW96} dark matter halo component, a MGE generated from a fit to a NFW profile can be added to the stellar component. [TO DO: Maybe remove detailed description of the mass MGE here, if it is described in detail somewhere else.] The gravitational potential is generated from the mass MGE by integrating the Poisson equation \citep{1994A&A...285..723E}. Equations \ref{eq:Jeans1} and \ref{eq:Jeans2} together with \ref{eq:bz} provide the velocity dispersion tensor $\langle v_i v_j \rangle = \langle v_i^2 \rangle$ (with $i,j \in \{ R,\phi, z\}$). It is then rotated by the inclination angle $i$ to the coordinate system of the observer ($(x',y')$ being the plane of the sky and $z'$ the line-of-sight, where $x'$ is alinged with the galaxy's major axis). Taking a light-weighted projection along the line-of-sight gives a model prediction for the line-of-sight velocity second moment, which is comparable to actual spectroscopic measurements of the second velocity moment. Details of the derivation using the MGE formalism are given in \citet{Cap08} and repeated in \citet{GlennEC}, we therefore just quote the result by \citet{Cap08} (their Equation 28) for the line-of-sight second velocity moment prediction from the Jeans equations,

@@

[TO DO: make primed and unprimed coordinate systems clear.]

\paragraph{Data comparison.} As data we use stellar line-of-sight rotation velocities $v_\text{rot} \equiv \langle v_\text{los} \rangle$ [TO DO: consistent, los or rot] and velocity dispersions $\sigma$ as described in Section \ref{sec:data}. The JAM models give us a prediction for the second line-of-sight velocity moment $\langle v_\text{los}^2 \rangle$. The root mean square (rms) line-of-sight velocity $v_\text{rms}$ allows a data-model comparison by relating theses velocities according to 
\begin{equation*}
v_\text{rms}^2 = \langle v_\text{los}^2 \rangle = v_\text{rot}^2 + \sigma^2.
\end{equation*}
The model in Equation \ref{eq:explicitLOSvelCap} predicts the intrinsic $\langle v_\text{los}^2 \rangle$ at a given position on the sky, which have then to be modified according to the mode of observation, to be comparable to the measurements. The measured $v_\text{rms}$ is a light-weighted mean for a pixel along the long-slit of the spectrograph, with height $L_y = 1$ arcsec \citep{SWELLSV} and a certain given extent in along the galaxy's major axis, $L_x$, i.e. for a rectangular aperture
\begin{equation*}
\text{AP}(x,y) = \left\{ \begin{array}{ll} 1 & \text{for } -\frac{L_x}{2} \leq x \leq + \frac{L_x}{2}\\
& \text{and } - \frac{L_y}{2} \leq y \leq + \frac{L_y}{2}  \\ 0 & \text{otherwise.} \end{array} \right.
\end{equation*}
The light arriving at the spectrograph itself was subject to seeing, i.e. a Gaussian with Full Width Half Maximum (FWHM) of1.1 arcsec \citep{SWELLSV}
\begin{equation*}
\text{PSF}(x,y)=\mathscr{N}(0,\text{FWHM}/2\sqrt{2\ln2}).
\end{equation*}
The model predictions have therefore to be convolved with the convolution kernel
\begin{eqnarray*}
K(x,y) &=& (\text{PSF} \ast \text{AP})(x,y) \\
&=& \frac{1}{4} \prod_{u \in \{x,y\}} \left[ \text{erf}\left( \frac{L_u/2 - u}{\sqrt{2}\sigma_\text{seeing}}\right) + \text{erf} \left( \frac{L_u/2 + u}{\sqrt{2} \sigma_\text{seeing}} \right) \right]
\end{eqnarray*}
and weighted by the surface brightness $I(x,y)$ [TO DO: primed x and y or not????]
\begin{eqnarray*}
I_\text{obs} &=& I \ast K\\
\langle v_\text{los}^2 \rangle |_\text{obs} &=& \frac{(I \langle v_\text{los}^2\rangle) \ast K}{I_\text{obs}}.
\end{eqnarray*}
If provided with the convolution kernel, the JAM code by \citet{Cap08} [TO DO: reference code] performs the convolution numerically. We set $L_x = 0.21$ arcsec as the width of the model pixel, and get a prediction for the actual measurements in bins of width 0.63, 1.26 and 1.89 arcsec \citep{SWELLSV} as light-weighted mean from each 3, 6 and 9 model pixels.

\paragraph{Rotation curve.} The intrinsic rotation curve is the first velocity moment $\langle v_\phi\rangle = \sqrt{\langle v_\phi^2 \rangle - \sigma_\phi^2}$. The observed rotation curve is the projection of the light-weighted contributions to $\langle v_\phi\rangle$ along the line-of-sight(\cite{Cap08}),
\begin{equation*}
I \langle v_\text{los}\rangle = \int_{-\infty}^{+\infty} \nu \langle v_\phi\rangle \cos \phi \sin i \diff z'.
\end{equation*}
The first velocity moments cannot be uniquely determined from the Jeans equations, which give only a prediction for the second velocity moments. Further assumptions are needed to separate the second velocity moments into ordered and random motion. \citet{Cap08} assumes that in a steady state there is no streaming velocity in $R$ direction, i.e. $\langle v_R \rangle = 0$ and therefore $\sigma_R^2 = \langle v_R^2 \rangle$. Then \citet{Cap08} relates the dispersions in $R$ and $\phi$ direction such that
$$\langle v_\phi\rangle = \sqrt{\langle v_\phi^2 \rangle - \sigma_\phi^2} \equiv \kappa \sqrt{\langle v_\phi^2 \rangle - \langle v_R^2 \rangle},$$
and the $\kappa$ parameter quantifies the rotation: $\kappa = 0$ means no rotation at all and $|\kappa| = 1$ describes a velocity dispersion ellipsoid that is a circle in the $R$-$\phi$ plane \citep{Cap08}. The sign of $\kappa$ determines the rotation direction. We can assign a constant $\kappa_k$ to every Gaussian in the MGE formalism and
\begin{equation*}
\nu \langle v_\phi\rangle = \left[\nu \sum_{k} \kappa_k^2 \left( [\nu\langle v_\phi^2 \rangle]_k - [\nu\langle v_R^2 \rangle]_k\right) \right]^{1/2}
\end{equation*} 
is then the light-weighted circular velocity curve, given the second velocity moments found from the Jeans equations.
\\To model the counter-rotating core of J1331 with one free parameter for, we employ the condition that the overall $\kappa(R)$ profile should smoothly and relatively steeply transition from $\kappa(R) = -\kappa' < 0$ at small $R$ [TO DO: Check, ich glaube das hier ist das intrinsische R in Zylinder-Coordinaten] trough $\kappa(R_0) \equiv 0$ and increase to $\kappa(R) = \kappa' > 0$ at large $R$. Our imposed profile is
\begin{equation}
\kappa(R) = \kappa' \frac{R^2 - R_0^2}{R^2 + R_0^2}. \label{eq:kappa_profile}
\end{equation}
We find $\kappa'$ by matching the model $\langle v_\text{los} \rangle$ with the symmetrized $v_\text{rot}$ data, where for a given $\kappa'$ the $\kappa_k$ are found from fitting the MGE generated profile $\kappa(R) = \sum_k \kappa_k \nu_k(r)/\sum_k \nu_k(r)$ to Equation \ref{eq:kappa_profile}. The observed zero point is at $R'_0\approx 2$ arcsec. In the deprojected galactic plane the radius of zero rotation would be at a $R_0 \gtrsim 2$ arcsec, and we choose it to be at 2.2 arcsec.

>>>>

\\The JAM modelling approach by \citet{Cap08} makes use of expressing the tracer density and the mass density as MGEs (see also \citet{1994A&A...285..723E}). The density of stellar tracers is assumed to be proportional to the observed and deprojected brightness distribution $\nu(R,z)$ in Equation \ref{eq:deprojMGE}. The mass density $\rho(R,z)$ can consist of several sets of MGEs, describing stellar and dark matter components. The gravitational potential is generated from the mass MGE by integrating the Poisson equation \citep{1994A&A...285..723E}. Equations \ref{eq:Jeans1} and \ref{eq:Jeans2} together with \ref{eq:bz} provide the velocity dispersion tensor $\langle v_i v_j \rangle = \langle v_i^2 \rangle$ (with $i,j \in \{ R,\phi, z\}$). It is then rotated by the inclination angle $i$ to the coordinate system of the observer ($(x',y')$ being the plane of the sky and $z'$ the line-of-sight, where $x'$ is alinged with the galaxy's major axis). Taking a light-weighted projection along the line-of-sight gives a model prediction for the line-of-sight velocity second moment, which is comparable to actual spectroscopic measurements of the second velocity moment. Details of the derivation using the MGE formalism are given in \citet{Cap08} and repeated in \citet{GlennEC}, we therefore just quote the result by \citet{Cap08} (their Equation 28) for the line-of-sight second velocity moment prediction from the Jeans equations,

@@

\paragraph{Data comparison.} As data we use stellar line-of-sight rotation velocities $v_\text{rot} \equiv \langle v_\text{los} \rangle$ [TO DO: consistent, los or rot] and velocity dispersions $\sigma$ as described in Section \ref{sec:data}. The JAM models give us a prediction for the second line-of-sight velocity moment $\langle v_\text{los}^2 \rangle$. The root mean square (rms) line-of-sight velocity $v_\text{rms}$ allows a data-model comparison by relating theses velocities according to 
\begin{equation*}
v_\text{rms}^2 = \langle v_\text{los}^2 \rangle = v_\text{rot}^2 + \sigma^2.
\end{equation*}
The model in Equation \ref{eq:explicitLOSvelCap} predicts the intrinsic $\langle v_\text{los}^2 \rangle$ at a given position on the sky, which have then to be modified according to the mode of observation, to be comparable to the measurements. The measured $v_\text{rms}$ is a light-weighted mean for a pixel along the long-slit of the spectrograph, with height $L_y = 1$ arcsec \citep{SWELLSV} and a certain given extent in along the galaxy's major axis, $L_x$, i.e. for a rectangular aperture
\begin{equation*}
\text{AP}(x',y') = \left\{ \begin{array}{ll} 1 & \text{for } -\frac{L_x}{2} \leq x' \leq + \frac{L_x}{2}\\
& \text{and } - \frac{L_y}{2} \leq y' \leq + \frac{L_y}{2}  \\ 0 & \text{otherwise.} \end{array} \right.
\end{equation*}
The light arriving at the spectrograph itself was subject to seeing, i.e. a Gaussian with Full Width Half Maximum (FWHM) of 1.1 arcsec \citep{SWELLSV}
\begin{equation*}
\text{PSF}(x',y')=\mathscr{N}(0,\text{FWHM}/2\sqrt{2\ln2}).
\end{equation*}
The model predictions have therefore to be convolved with the convolution kernel
\begin{eqnarray*}
K(x',y') &=& (\text{PSF} \ast \text{AP})(x',y') \\
&=& \frac{1}{4} \prod_{u \in \{x',y'\}} \left[ \text{erf}\left( \frac{L_u/2 - u}{\sqrt{2}\sigma_\text{seeing}}\right) + \text{erf} \left( \frac{L_u/2 + u}{\sqrt{2} \sigma_\text{seeing}} \right) \right]
\end{eqnarray*}
and weighted by the surface brightness $I(x',y')$
\begin{eqnarray*}
I_\text{obs} &=& I \ast K\\
\langle v_\text{los}^2 \rangle |_\text{obs} &=& \frac{(I \langle v_\text{los}^2\rangle) \ast K}{I_\text{obs}}.
\end{eqnarray*}
If provided with the convolution kernel, the JAM code by \citet{Cap08} [TO DO: reference code] performs the convolution numerically. We set $L_x = 0.21$ arcsec as the width of the model pixel, and get a prediction for the actual measurements in bins of width 0.63, 1.26 and 1.89 arcsec \citep{SWELLSV} as light-weighted mean from each 3, 6 and 9 model pixels.

\paragraph{Rotation curve.} The intrinsic rotation curve is the first velocity moment $\langle v_\phi\rangle = \sqrt{\langle v_\phi^2 \rangle - \sigma_\phi^2}$. The observed rotation curve is the projection of the light-weighted contributions to $\langle v_\phi\rangle$ along the line-of-sight(\cite{Cap08}),
\begin{equation*}
I \langle v_\text{los}\rangle = \int_{-\infty}^{+\infty} \nu \langle v_\phi\rangle \cos \phi \sin i \diff z'.
\end{equation*}
The first velocity moments cannot be uniquely determined from the Jeans equations, which give only a prediction for the second velocity moments. Further assumptions are needed to separate the second velocity moments into ordered and random motion. \citet{Cap08} assumes that in a steady state there is no streaming velocity in $R$ direction, i.e. $\langle v_R \rangle = 0$ and therefore $\sigma_R^2 = \langle v_R^2 \rangle$. Then \citet{Cap08} relates the dispersions in $R$ and $\phi$ direction such that
$$\langle v_\phi\rangle = \sqrt{\langle v_\phi^2 \rangle - \sigma_\phi^2} \equiv \kappa \sqrt{\langle v_\phi^2 \rangle - \langle v_R^2 \rangle},$$
and the $\kappa$ parameter quantifies the rotation: $\kappa = 0$ means no rotation at all and $|\kappa| = 1$ describes a velocity dispersion ellipsoid that is a circle in the $R$-$\phi$ plane \citep{Cap08}. The sign of $\kappa$ determines the rotation direction. We can assign a constant $\kappa_k$ to every Gaussian in the MGE formalism and
\begin{equation*}
\nu \langle v_\phi\rangle = \left[\nu \sum_{k} \kappa_k^2 \left( [\nu\langle v_\phi^2 \rangle]_k - [\nu\langle v_R^2 \rangle]_k\right) \right]^{1/2}
\end{equation*} 
is then the light-weighted circular velocity curve, given the second velocity moments found from the Jeans equations.
\\To model the counter-rotating core of J1331 with one free parameter for, we employ the condition that the overall $\kappa(R)$ profile should smoothly and relatively steeply transition from $\kappa(R) = -\kappa' < 0$ at small $R$ through $\kappa(R_0) \equiv 0$ and increase to $\kappa(R) = \kappa' > 0$ at large $R$. Our imposed profile is
\begin{equation}
\kappa(R) = \kappa' \frac{R^2 - R_0^2}{R^2 + R_0^2}. \label{eq:kappa_profile}
\end{equation}
We find $\kappa'$ by matching the model $\langle v_\text{los} \rangle$ with the symmetrized $v_\text{rot}$ data, where for a given $\kappa'$ the $\kappa_k$ are found from fitting the MGE generated profile $\kappa(R) = \sum_k \kappa_k \nu_k(r)/\sum_k \nu_k(r)$ to Equation \ref{eq:kappa_profile}. The observed zero point is at $R'_0\approx 2$ arcsec. In the deprojected galactic plane the radius of zero rotation would be at a $R_0 \gtrsim 2$ arcsec, and we choose it to be at 2.2 arcsec.

### 2015-09-11 docs/model_MGE_v1.01.tex
Removed some TO DOs.

Multi-Gaussian Expansions (MGE) are used to parametrize the observed surface brightness or projected total mass of a galaxy as a sum of $N$ two-dimensional, elliptical Gaussians \citep{1991ApJ...366..599B,1992A&A...253..366M,1994A&A...285..723E,1999MNRAS.303..495E}. This work makes use of the algorithm and code\footnote{Michele Cappellari's IDL code package for fitting MGEs to images is available online at \url{http://www-astro.physics.ox.ac.uk/~mxc/software}. The version from June 2012 was used in this work.} by \citet{Cap02}. We assume all Gaussians to have the same center and position angle $\phi$, i.e. orientation of w.r.t. the $y'$-axis of the coordinate system with polar coordinates $(R',\theta')$ [TO DO: CHECK]. Then the surface brightness can be written as
\begin{eqnarray}
I(R',\theta') &=& \sum_{i=1}^{N} I_{0,i} \exp\left[ - \frac{1}{2\sigma_i^2} \left(x^{'2} + \frac{y^{'2}}{q_i^{'2}}\right)\right]\label{eq:MGEgeneral}\\
\text{with } I_{0,i} &=& \frac{L_i}{2\pi \sigma_i^2 q'_i}\label{eq:centralItotalL}\\
\text{and } x'_i &=& R' \cos(\theta' - \phi)\nonumber\\
y'_i &=& R' \sin(\theta' - \phi),\nonumber
\end{eqnarray}
where $I_{0,i}$ is the central surface brightness of each Gaussian, $L_i$ its total luminosity, $\sigma_i$ its dispersion along the major axis and $q'_i$ the axis ratio between the elliptical Gaussians major and minor axis.
\\We can also expand the telescopes point-spread function (PSF) as a sum of circular Gaussians,
\begin{equation}
\text{PSF}(x,y) = \sum_j \frac{G_j}{2 \pi \delta_j^2} \exp\left[- \frac{1}{2 \delta_j^2} \left(x^2 + y^2 \right)\right], \label{eq:PSFgeneral}
\end{equation}
where $\sum_j G_j = 1$ and $\delta_j$ are in this case the dispersions of the circular PSF Gaussians. In this case the observed surface brightness distribution is a convolution of the intrinsic surface brightness in Eq. (\ref{eq:MGEgeneral}) with the PSF in Eq. (\ref{eq:PSFgeneral}): $(I \ast \text{PSF}) (x',y')$ is then again a sum of Gaussians and can be directly fitted to an image of the galaxy in question.

>>>>

Multi-Gaussian Expansions (MGE) are used to parametrize the observed surface brightness or projected total mass of a galaxy as a sum of $N$ two-dimensional, elliptical Gaussians \citep{1991ApJ...366..599B,1992A&A...253..366M,1994A&A...285..723E,1999MNRAS.303..495E}. This work makes use of the algorithm and code\footnote{Michele Cappellari's IDL code package for fitting MGEs to images is available online at \url{http://www-astro.physics.ox.ac.uk/~mxc/software}. The version from June 2012 was used in this work.} by \citet{Cap02}. We assume all Gaussians to have the same center and position angle $\phi$, i.e. orientation of the Gaussians' major axis measured counter-clockwise from the $y'$-axis of the coordinate system with polar coordinates $(R',\theta')$. Then the surface brightness can be written as
\begin{eqnarray}
I(R',\theta') &=& \sum_{i=1}^{N} I_{0,i} \exp\left[ - \frac{1}{2\sigma_i^2} \left(x^{'2} + \frac{y^{'2}}{q_i^{'2}}\right)\right]\label{eq:MGEgeneral}\\
\text{with } I_{0,i} &=& \frac{L_i}{2\pi \sigma_i^2 q'_i}\label{eq:centralItotalL}\\
\text{and } x'_i &=& R' \cos(\theta' - \phi)\nonumber\\
y'_i &=& R' \sin(\theta' - \phi),\nonumber
\end{eqnarray}
where $I_{0,i}$ is the central surface brightness of each Gaussian, $L_i$ its total luminosity, $\sigma_i$ its dispersion along the major axis and $q'_i$ the axis ratio between the elliptical Gaussians major and minor axis.
\\We can also expand the telescopes point-spread function (PSF) as a sum of circular Gaussians,
\begin{equation}
\text{PSF}(x',y') = \sum_j \frac{G_j}{2 \pi \delta_j^2} \exp\left[- \frac{1}{2 \delta_j^2} \left({x'}^2 + {y'}^2 \right)\right], \label{eq:PSFgeneral}
\end{equation}
where $\sum_j G_j = 1$ and $\delta_j$ are in this case the dispersions of the circular PSF Gaussians. In this case the observed surface brightness distribution is a convolution of the intrinsic surface brightness in Equation \ref{eq:MGEgeneral} with the PSF in Equation \ref{eq:PSFgeneral}: $(I \ast \text{PSF}) (x',y')$ is then again a sum of Gaussians and can be directly fitted to an image of the galaxy in question.
